\papertitle

### 2013-02-01 manuscript.tex
fixed figure references

To explore these cooperative control concepts, we designed a new 
flapping-wing MAV, known as the H$^2$Bird (Fig.~\ref{fig:h2bird}). Built 
around the Silverlit i-Bird RC flyer power train and clap-fling wings\footnote{\raggedright Silverlit Toys Manufactory Ltd.: i-Bird RC Flyer
\href{http://www.silverlit-flyingclub.com/wingsmaster/}
     {http://www.silverlit-flyingclub.com/wingsmaster/}}, 
the H$^2$Bird has a wingspan of 26.5 cm and a flight weight of 13 grams. A 
tail propeller and servo-controlled tail elevator provide control in the yaw 
and pitch axes (Fig.~\ref{fig:h2Bird_axes}). The on-board ImageProc 2.4\footnote{ImageProc 2.4: \\
\href{https://github.com/biomimetics/imageproc\_pcb}
     {https://github.com/biomimetics/imageproc\_pcb}} 
controller includes of a 40 MIPS microprocessor, 6 DOF IMU, 
IEEE 802.15.4 radio, VGA camera, and motor drivers, all powered by a 90 mAh 
lithium polymer battery. In routine flight, the H$^2$Bird averages 1.2 m/s 
ground speed, and operates for approximately 10 minutes.

@@

We implement attitude estimation and control of the H$^2$Bird on-board to 
achieve high sample-rate attitude control. We estimate the vehicle pose by
integrating the IMU, and use a proportional-integral-derivative (PID) 
controller to feed these estimates back to the attitude control flight 
surfaces (Figs.~\ref{fig:process_flow} and~\ref{fig:block_diagram}).

@@

Our implementation of quaternion-based control is similar to 
Knoebe's~\cite{knoebe:quatcontrol}:
Given a reference pose represented in quaternion form $q_r$, we define the 
error as the rotation $q_e$ required to reach the reference pose from the 
estimated pose $q$:
\begin{align}
\label{eq:quat_error}
q_e &= q^{\prime}\cdot q_r \\
		&= q_{e,w} + q_{e,x}\hat{i} + q_{e,y}\hat{j} + q_{e,z}\hat{k} \notag 
\end{align}
We then convert this error quaternion to an angle axis representation and 
use it as input to the PID controller (Eqs.~\ref{eq:quat_linearize} and~\ref{eq:quat_angle}).
We perform these calculations on-board, with trigonometric lookup tables for 
real-time performance.
\begin{align}
\label{eq:quat_linearize}
Q_e &= \alpha\cdot q_e/\sin(\alpha /2) \\
		&= \alpha_{x}\hat{i} + \alpha_{y}\hat{j} + \alpha_{z}\hat{k} \notag
\end{align}
\begin{equation}
\label{eq:quat_angle}
\alpha = 2\arccos(q_{e,w})
\end{equation}

%----------------------------------------------------------------------------%
\subsection{Ground Station Pose Estimation}
The ground station performs simple two-dimensional pose estimation on 
H$^2$Bird using a particle filter-based motion tracking algorithm~\cite{thrun2005probabilistic}. This pose 
estimate is the feedback input to the cooperative visual servoing feedback 
loop, which is represented as the outermost feedback paths on the block 
diagrams in Figs.~\ref{fig:process_flow} and~\ref{fig:block_diagram}. We 
limit our pose estimation approach to monocular two-dimensional object 
tracking in pixel space. This allows us to perform real-time video tracking 
using the modest computational resources available on the ground station, 
and similarly, millirobotic ground vehicles.

We initialize the particles 
uniformly across the frame before tracking begins. To improve numerical 
stability, we normalize the weights of all particles on each iteration, and 
only resample the particle population when the mean particle weight falls 
below a pre-determined threshold. 
%............................................................................%
\subsubsection{Motion Model}
Our motion model (Eq.~\ref{eq:motion_model}) is a simple Gaussian transition 
centered around each particle's current position, augmented with an 
$\epsilon$-random uniform sampling strategy. In Eq.~\ref{eq:motion_model}, 
$x^{[i]}_t$ is the position of the $i^\text{th}$ particle in pixel space at 
time $t$, $\Sigma$ is a chosen covariance matrix, and $b$ is a vector of the 
bounds of the pixel space.
\begin{equation}
\label{eq:motion_model}
x^{[i]}_t \sim \begin{cases}
\mathcal{N}(x^{[i]}_{t-1}, \Sigma), &\text{with probability } 1-\epsilon \\
\mathcal{U}(0,b),& \text{with probability } \epsilon \\
\end{cases}\\
\end{equation}
In the experiments we present in this work, we used a diagonal covariance matrix 
with a variance of 256, and an $\epsilon$ of 0.3. We find the Gaussian 
transition model is sufficient for tracking motion in which frame-to-frame 
position changes are relatively small. The $\epsilon$-random uniform 
transitions increase tracking robustness, by forcing the filter to always 
sample broadly across the pixel space. This helps prevent overconfidence 
and collapse, and decreases reacquisition delay.

%............................................................................%
\subsubsection{Emission Model}
Our emission model (Eq.~\ref{eq:emission_model}) implements a sampling
approach to motion tracking via na\"{i}ve background subtraction, e.g. frame 
differencing~\cite{Ahad2011Computer}. On each iteration, the system provides the emission 
model with $f_0$, the frame representing the background, and $f_t$, the most 
recent frame captured. The model assigns a weight $w^{[i]}_t$ to each 
particle $x^{[i]}_t$, proportional to the color distance in RGB space 
between the pixels of $f_0$ and $f_t$ at the location $x^{[i]}_t$.
\begin{equation}
\label{eq:emission_model}
w^{[i]}_t = 1 + \norm{f_0(x^{[i]}_t)-f_t(x^{[i]}_t)}^2
\end{equation} 
Hence, particles are more likely to be resampled in regions of 
the image which are dissimilar in color from the background.

@@

%----------------------------------------------------------------------------%
\subsection{Cooperative Visual Servoing}
\label{sec:visual_servoing_concept}
The ground station guides the H$^2$Bird by setting the
reference attitude of its on-board attitude 
controller (Figs.\ref{fig:process_flow} and~\ref{fig:block_diagram}). Using 
the estimate of 2D vehicle pose from the pose estimation algorithm, the 
ground station calculates an altitude error and horizontal error. The ground 
station then uses a PID control law to calculate a new reference pitch and yaw, 
which it sends to the H$^2$Bird over radio. 
We apply a fixed-lag window to the PID 
controller integral term to prevent wind-up while we launch the ornithopter. 
We tuned the proportional and derivative terms of this outer controller to 
dampen the overshoot and drift caused by the cooperative system latency.
\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/block_diagrams.pdf}
\caption{Overall block diagram of the cooperative control system.}
\label{fig:block_diagram}
\end{figure}

@@

%............................................................................%
\subsubsection{Ornithopter Attitude Control}
\label{sec:flight_control}
We determined the turning speed and turning radius of the H$^2$Bird 
experimentally, by measuring the response of the system to a step input of a 
clockwise 90$^{\circ}$ turn. We measured the response of the system by 
recording the attitude estimates calculated by the ornithopter using its 
on-board gyroscope. We then used MATLAB\footnote{The MathWorks, Inc. Matlab:
\href{http://www.mathworks.com/products/matlab/}
     {http://www.mathworks.com/products/matlab/}} 
to fit a simple low-order model to the step response, as discussed in 
Sec.~\ref{sec:model_attitude}.

@@

Each trial included the following steps (see Fig.~\ref{fig:pf_screencap}):
\begin{spaced_description}{\textbf{(4-6)}}
\item[\textbf{(1,2)}] We identify the window to the ground station by dragging a bounding 
box over it using a mouse. 
\item[\textbf{(3)}] We release H$^2$Bird by hand, in view of the camera at the desired 
starting grid point.
\item[\textbf{(4-6)}] The H$^2$Bird attempts to fly through the window, with remote 
guidance from the ground station.
\end{spaced_description}

>>>>

To explore these cooperative control concepts, we designed a new 
flapping-wing MAV, known as the H$^2$Bird (Figure~\ref{fig:h2bird}). Built 
around the Silverlit i-Bird RC flyer power train and clap-fling wings\footnote{\raggedright Silverlit Toys Manufactory Ltd.: i-Bird RC Flyer
\href{http://www.silverlit-flyingclub.com/wingsmaster/}
     {http://www.silverlit-flyingclub.com/wingsmaster/}}, 
the H$^2$Bird has a wingspan of 26.5 cm and a flight weight of 13 grams. A 
tail propeller and servo-controlled tail elevator provide control in the yaw 
and pitch axes (Figure~\ref{fig:h2Bird_axes}). The on-board ImageProc 2.4\footnote{ImageProc 2.4: \\
\href{https://github.com/biomimetics/imageproc\_pcb}
     {https://github.com/biomimetics/imageproc\_pcb}} 
controller includes of a 40 MIPS microprocessor, 6 DOF IMU, 
IEEE 802.15.4 radio, VGA camera, and motor drivers, all powered by a 90 mAh 
lithium polymer battery. In routine flight, the H$^2$Bird averages 1.2 m/s 
ground speed, and operates for approximately 10 minutes.

@@

We implement attitude estimation and control of the H$^2$Bird on-board to 
achieve high sample-rate attitude control. We estimate the vehicle pose by
integrating the IMU, and use a proportional-integral-derivative (PID) 
controller to feed these estimates back to the attitude control flight 
surfaces (Figures~\ref{fig:process_flow} and~\ref{fig:block_diagram}).

@@

Our implementation of quaternion-based control is similar to 
Knoebe's~\cite{knoebe:quatcontrol}:
Given a reference pose represented in quaternion form $q_r$, we define the 
error as the rotation $q_e$ required to reach the reference pose from the 
estimated pose $q$:
\begin{align}
\label{eq:quat_error}
q_e &= q^{\prime}\cdot q_r \\
		&= q_{e,w} + q_{e,x}\hat{i} + q_{e,y}\hat{j} + q_{e,z}\hat{k} \notag 
\end{align}
We then convert this error quaternion to an angle axis representation and 
use it as input to the PID controller (Equations~\ref{eq:quat_linearize} and~\ref{eq:quat_angle}).
We perform these calculations on-board, with trigonometric lookup tables for 
real-time performance.
\begin{align}
\label{eq:quat_linearize}
Q_e &= \alpha\cdot q_e/\sin(\alpha /2) \\
		&= \alpha_{x}\hat{i} + \alpha_{y}\hat{j} + \alpha_{z}\hat{k} \notag
\end{align}
\begin{equation}
\label{eq:quat_angle}
\alpha = 2\arccos(q_{e,w})
\end{equation}

%----------------------------------------------------------------------------%
\subsection{Ground Station Pose Estimation}
The ground station performs simple two-dimensional pose estimation on 
H$^2$Bird using a particle filter-based motion tracking algorithm~\cite{thrun2005probabilistic}. This pose 
estimate is the feedback input to the cooperative visual servoing feedback 
loop, which is represented as the outermost feedback paths on the block 
diagrams in Figures~\ref{fig:process_flow} and~\ref{fig:block_diagram}. We 
limit our pose estimation approach to monocular two-dimensional object 
tracking in pixel space. This allows us to perform real-time video tracking 
using the modest computational resources available on the ground station, 
and similarly, millirobotic ground vehicles.

We initialize the particles 
uniformly across the frame before tracking begins. To improve numerical 
stability, we normalize the weights of all particles on each iteration, and 
only resample the particle population when the mean particle weight falls 
below a pre-determined threshold. 
%............................................................................%
\subsubsection{Motion Model}
Our motion model (Equation~\ref{eq:motion_model}) is a simple Gaussian transition 
centered around each particle's current position, augmented with an 
$\epsilon$-random uniform sampling strategy. In Equation~\ref{eq:motion_model}, 
$x^{[i]}_t$ is the position of the $i^\text{th}$ particle in pixel space at 
time $t$, $\Sigma$ is a chosen covariance matrix, and $b$ is a vector of the 
bounds of the pixel space.
\begin{equation}
\label{eq:motion_model}
x^{[i]}_t \sim \begin{cases}
\mathcal{N}(x^{[i]}_{t-1}, \Sigma), &\text{with probability } 1-\epsilon \\
\mathcal{U}(0,b),& \text{with probability } \epsilon \\
\end{cases}\\
\end{equation}
In the experiments we present in this work, we used a diagonal covariance matrix 
with a variance of 256, and an $\epsilon$ of 0.3. We find the Gaussian 
transition model is sufficient for tracking motion in which frame-to-frame 
position changes are relatively small. The $\epsilon$-random uniform 
transitions increase tracking robustness, by forcing the filter to always 
sample broadly across the pixel space. This helps prevent overconfidence 
and collapse, and decreases reacquisition delay.

%............................................................................%
\subsubsection{Emission Model}
Our emission model (Equation~\ref{eq:emission_model}) implements a sampling
approach to motion tracking via na\"{i}ve background subtraction, e.g. frame 
differencing~\cite{Ahad2011Computer}. On each iteration, the system provides the emission 
model with $f_0$, the frame representing the background, and $f_t$, the most 
recent frame captured. The model assigns a weight $w^{[i]}_t$ to each 
particle $x^{[i]}_t$, proportional to the color distance in RGB space 
between the pixels of $f_0$ and $f_t$ at the location $x^{[i]}_t$.
\begin{equation}
\label{eq:emission_model}
w^{[i]}_t = 1 + \norm{f_0(x^{[i]}_t)-f_t(x^{[i]}_t)}^2
\end{equation} 
Hence, particles are more likely to be resampled in regions of 
the image which are dissimilar in color from the background.

@@

%----------------------------------------------------------------------------%
\subsection{Cooperative Visual Servoing}
\label{sec:visual_servoing_concept}
The ground station guides the H$^2$Bird by setting the
reference attitude of its on-board attitude 
controller (Figures~\ref{fig:process_flow} and~\ref{fig:block_diagram}). Using 
the estimate of 2D vehicle pose from the pose estimation algorithm, the 
ground station calculates an altitude error and horizontal error. The ground 
station then uses a PID control law to calculate a new reference pitch and yaw, 
which it sends to the H$^2$Bird over radio. 
We apply a fixed-lag window to the PID 
controller integral term to prevent wind-up while we launch the ornithopter. 
We tuned the proportional and derivative terms of this outer controller to 
dampen the overshoot and drift caused by the cooperative system latency.
\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/block_diagrams.pdf}
\caption{Overall block diagram of the cooperative control system.}
\label{fig:block_diagram}
\end{figure}

@@

%............................................................................%
\subsubsection{Ornithopter Attitude Control}
\label{sec:flight_control}
We determined the turning speed and turning radius of the H$^2$Bird 
experimentally, by measuring the response of the system to a step input of a 
clockwise 90$^{\circ}$ turn. We measured the response of the system by 
recording the attitude estimates calculated by the ornithopter using its 
on-board gyroscope. We then used MATLAB\footnote{The MathWorks, Inc. Matlab:
\href{http://www.mathworks.com/products/matlab/}
     {http://www.mathworks.com/products/matlab/}} 
to fit a simple low-order model to the step response, as discussed in 
Section~\ref{sec:model_attitude}.

@@

Each trial included the following steps (see Figure~\ref{fig:pf_screencap}):
\begin{spaced_description}{\textbf{(4-6)}}
\item[\textbf{(1,2)}] We identify the window to the ground station by dragging a bounding 
box over it using a mouse. 
\item[\textbf{(3)}] We release H$^2$Bird by hand, in view of the camera at the desired 
starting grid point.
\item[\textbf{(4-6)}] The H$^2$Bird attempts to fly through the window, with remote 
guidance from the ground station.
\end{spaced_description}
